---
title: "Denominator neglect data tests for learning effects"
author: "Sándor Máté Csaba"
date: "2/3/2022"
output: html_document
---
## Learning effect rtests on Szaszi et al

### Varaible map

```{r, message = FALSE, warning = FALSE}
library(tidyverse)
readxl::read_excel("szaszi_sup/Description of the variables.xlsx")
```

### Experiment 1 data

```{r, message = FALSE, warning = FALSE}
exp1_data <- readr::read_delim("szaszi_sup/exp1_data.csv", 
    ";", escape_double = FALSE, trim_ws = TRUE)

summary(exp1_data)
```

### Experiment 2 data

```{r, message = FALSE, warning = FALSE}
exp2_data <- readr::read_delim("szaszi_sup/exp2_data.csv", 
    ";", escape_double = FALSE, trim_ws = TRUE)
summary(exp2_data)
```

## Experiment 2 raw data, testing if the mt package keeps the experiment order (spoiler alert, it does)

```{r, message = FALSE, warning = FALSE}
raw_data_exp2 <- read_delim("szaszi_sup/Raw_data exp2.csv", 
    ";", escape_double = FALSE, trim_ws = TRUE)

raw_data_exp2

exp2_data %>% 
  select(ID, RT) %>%
  group_by(ID) %>%
  slice_head(n = 3) %>%
  mutate(row_number = row_number()) %>%
  ungroup() %>%
  left_join(raw_data_exp2 %>% 
  select(ID, RT) %>%
  group_by(ID) %>%
  slice_head(n = 3) %>%
  mutate(row_number = row_number()) %>%
  ungroup(), by = c("ID","row_number")) %>%
  mutate(diff = RT.x-RT.y)


sum(raw_data_exp2$Left == raw_data_exp2$Right)
```

### Success rate vs trial no. controlled for congruency / first-last / numeracy / intelligence

```{r, message = FALSE, warning = FALSE}
#choose dataset
szaszi_data <- exp1_data

#only congruency
szaszi_data %>%
  mutate(IQ_cat = ifelse(IQ < 119,"Low","High")) %>%
  mutate(numeracy_cat = ifelse(numeracy < 3,"Low","High")) %>%
  mutate(outcome_first = Correct & !(CoM %% 2)) %>%
  select(ID, mt_id, cong = Congruency_f, outcome_last = Correct, outcome_first, CoM, numeracy_cat, IQ_cat) %>%
  pivot_longer(cols = c("outcome_first", "outcome_last"),
               names_to = "outcome_type",
               values_to = "outcome_value") %>%
  group_by(ID) %>%
  mutate(trial_no = seq_along(mt_id)) %>%
  ungroup()  %>%
  group_by(trial_no, cong, outcome_type) %>%
  summarize(suc_rate = sum(outcome_value)/n(), .groups = "drop") %>%
  {ggplot(.,aes(x = trial_no, y = suc_rate, fill = cong)) +
    #geom_bar(stat="identity", position=position_dodge()) +
    geom_smooth(method = "loess") +
    ggtitle("Success rate vs. trial no. at congruency and first-last outcome") +
    #coord_cartesian(ylim = c(min(.$suc_rate)*.75, 1)) +
    theme_minimal() +
    facet_grid(rows = vars(outcome_type))}

#numeracy
szaszi_data %>%
  mutate(IQ_cat = ifelse(IQ < 119,"Low","High")) %>%
  mutate(numeracy_cat = ifelse(numeracy < 3,"Low","High")) %>%
  mutate(outcome_first = Correct & !(CoM %% 2)) %>%
  select(ID, mt_id, cong = Congruency_f, outcome_last = Correct, outcome_first, CoM, numeracy_cat, IQ_cat) %>%
  pivot_longer(cols = c("outcome_first", "outcome_last"),
               names_to = "outcome_type",
               values_to = "outcome_value") %>%
  group_by(ID) %>%
  mutate(trial_no = seq_along(mt_id)) %>%
  ungroup()  %>%
  group_by(trial_no, cong, outcome_type, numeracy_cat) %>%
  summarize(suc_rate = sum(outcome_value)/n(), .groups = "drop") %>%
  {ggplot(.,aes(x = trial_no, y = suc_rate, fill = cong)) +
    #geom_bar(stat="identity", position=position_dodge()) +
    geom_smooth(method = "loess") +
    ggtitle("Success rate vs. trial no. at congruency and first-last outcome and numeracy") +
    #coord_cartesian(ylim = c(min(.$suc_rate)*.75, 1)) +
    theme_minimal() +
    facet_grid(rows = vars(outcome_type), cols = vars(numeracy_cat))}

#IQ
szaszi_data %>%
  mutate(IQ_cat = ifelse(IQ < 119,"Low","High")) %>%
  mutate(numeracy_cat = ifelse(numeracy < 3,"Low","High")) %>%
  mutate(outcome_first = Correct & !(CoM %% 2)) %>%
  select(ID, mt_id, cong = Congruency_f, outcome_last = Correct, outcome_first, CoM, numeracy_cat, IQ_cat) %>%
  pivot_longer(cols = c("outcome_first", "outcome_last"),
               names_to = "outcome_type",
               values_to = "outcome_value") %>%
  group_by(ID) %>%
  mutate(trial_no = seq_along(mt_id)) %>%
  ungroup()  %>%
  group_by(trial_no, cong, outcome_type, IQ_cat) %>%
  summarize(suc_rate = sum(outcome_value)/n(), .groups = "drop") %>%
  {ggplot(.,aes(x = trial_no, y = suc_rate, fill = cong)) +
    #geom_bar(stat="identity", position=position_dodge()) +
    geom_smooth(method = "loess") +
    ggtitle("Success rate vs. trial no. at congruency and first-last outcome and IQ") +
    coord_cartesian(ylim = c(min(.$suc_rate)*.75, 1)) +
    theme_minimal() +
    facet_grid(rows = vars(outcome_type), cols = vars(IQ_cat))}
```

### Change of Mind rate vs trial no. controlled for congruency / first-last / numeracy / intelligence

```{r, message = FALSE, warning = FALSE}
#choose dataset
szaszi_data <- exp1_data

#only congruency
  
szaszi_data %>%
  mutate(IQ_cat = ifelse(IQ < 119,"Low","High")) %>%
  mutate(numeracy_cat = ifelse(numeracy < 3,"Low","High")) %>%
  select(ID, mt_id, cong = Congruency_f, CoM, IQ_cat, numeracy_cat) %>%
  group_by(ID) %>%
  mutate(trial_no = seq_along(mt_id)) %>%
  ungroup() %>%
  group_by(trial_no, cong) %>%
  summarize(inst_rate = sum(CoM)/n()) %>%
  {ggplot(.,aes(x = trial_no, y = inst_rate, fill = cong)) +
    #geom_bar(stat="identity", position=position_dodge()) +
    geom_smooth(method = "loess") +
    ggtitle("Change of Mind rate vs. trial no. at congruency") +
    #coord_cartesian(ylim = c(min(.$suc_rate)*.75, 1)) +
    #facet_grid(cols = vars(IQ_cat)) +
    theme_minimal()}

#numeracy
  
szaszi_data %>%
  mutate(IQ_cat = ifelse(IQ < 119,"Low","High")) %>%
  mutate(numeracy_cat = ifelse(numeracy < 3,"Low","High")) %>%
  select(ID, mt_id, cong = Congruency_f, CoM, IQ_cat, numeracy_cat) %>%
  group_by(ID) %>%
  mutate(trial_no = seq_along(mt_id)) %>%
  ungroup() %>%
  group_by(trial_no, cong, numeracy_cat) %>%
  summarize(inst_rate = sum(CoM)/n()) %>%
  {ggplot(.,aes(x = trial_no, y = inst_rate, fill = cong)) +
    #geom_bar(stat="identity", position=position_dodge()) +
    geom_smooth(method = "loess") +
    ggtitle("Change of Mind rate vs. trial no. at congruency and numeracy") +
    #coord_cartesian(ylim = c(min(.$suc_rate)*.75, 1)) +
    facet_grid(cols = vars(numeracy_cat)) +
    theme_minimal()}

#IQ
  
exp1_data %>%
  mutate(IQ_cat = ifelse(IQ < 119,"Low","High")) %>%
  mutate(numeracy_cat = ifelse(numeracy < 3,"Low","High")) %>%
  select(ID, mt_id, cong = Congruency_f, CoM, IQ_cat, numeracy_cat) %>%
  group_by(ID) %>%
  mutate(trial_no = seq_along(mt_id)) %>%
  ungroup() %>%
  group_by(trial_no, cong, IQ_cat) %>%
  summarize(inst_rate = sum(CoM)/n()) %>%
  {ggplot(.,aes(x = trial_no, y = inst_rate, fill = cong)) +
    #geom_bar(stat="identity", position=position_dodge()) +
    geom_smooth(method = "loess") +
    ggtitle("Change of Mind rate vs. trial no. at congruency and IQ") +
    #coord_cartesian(ylim = c(min(.$suc_rate)*.75, 1)) +
    facet_grid(cols = vars(IQ_cat)) +
    theme_minimal()}

```

## Learning effect rtests on Szaszi et al

### Load datasets
```{r, message = FALSE, warning = FALSE}
experiment_data <- read_delim("barna_sup/subject-1.csv", ",", escape_double = FALSE, trim_ws = TRUE, col_types = cols()) %>% 
  mutate(ID = 1)
for(i in seq(2,185)){
  fname <- paste0("barna_sup/subject-",i,".csv")
  if(file.exists(fname)){
    experiment_data <- experiment_data %>%
      bind_rows(
        read_delim(fname, ",", escape_double = FALSE, trim_ws = TRUE, col_types = cols()) %>% 
          mutate(ID = i)
      )
  }
}

pic_to_value <- function(x){
  if(substr(x,1,1) == "s"){
    return(substr(x,2,2) %>% as.numeric() %>% {./10})
  }else{
    return(x  %>% as.numeric() %>% {./100})
  }
}

pic_to_base <- function(x){
  if(substr(x,1,1) == "s"){
    return("s")
  }else{
    return("b")
  }
}

formated_experiment_data <- experiment_data %>%
  drop_na(any_of(c("Button_response","Left","Right"))) %>%
  mutate(left_value = map(Left, pic_to_value) %>% unlist(),
         left_c = map(Left, pic_to_base) %>% unlist(),
         right_value = map(Right, pic_to_value) %>% unlist(),
         right_c = map(Right, pic_to_base) %>% unlist(),
         response_value = map(Button_response, pic_to_value) %>% unlist()) %>%
  mutate(correct = response_value == pmax(left_value, right_value),
         congruency = ifelse(left_value > right_value,
                             ifelse(left_c == "s", "Incongruent", "Congruent"),
                             ifelse(right_c == "s", "Incongruent", "Congruent")
                             )
         )
```
### Success rate vs trial no. controlled for instuction / gender

```{r}
#basic
formated_experiment_data %>%
  mutate(instruction = ifelse(is.na(instruction),"Not","Instructed")) %>%
  select(ID, live_row, correct, congruency, instruction, income_level, gender, handedness) %>%
  drop_na() %>%
  group_by(live_row, congruency) %>%
  summarize(suc_rate = sum(correct)/n(), .groups = "drop") %>%
  {ggplot(.,aes(x = live_row, y = suc_rate, fill = congruency)) +
    #geom_bar(stat="identity", position=position_dodge()) +
    geom_smooth(method = "loess") +
    ggtitle("Success rate vs. trial no. at congruency") +
    #coord_cartesian(ylim = c(min(.$suc_rate)*.75, 1)) +
    #facet_grid(rows = vars(outcome_type)) +
    theme_minimal()}
#basic
formated_experiment_data %>%
  mutate(instruction = ifelse(is.na(instruction),"Not","Instructed")) %>%
  select(ID, live_row, correct, congruency, instruction, income_level, gender, handedness) %>%
  drop_na() %>%
  group_by(live_row, congruency, instruction) %>%
  summarize(suc_rate = sum(correct)/n(), .groups = "drop") %>%
  {ggplot(.,aes(x = live_row, y = suc_rate, fill = congruency)) +
    #geom_bar(stat="identity", position=position_dodge()) +
    geom_smooth(method = "loess") +
    ggtitle("Success rate vs. trial no. at instruction") +
    #coord_cartesian(ylim = c(min(.$suc_rate)*.75, 1)) +
    facet_grid(rows = vars(instruction)) +
    theme_minimal()}

#gender
formated_experiment_data %>%
  mutate(instruction = ifelse(is.na(instruction),"Not","Instructed")) %>%
  select(ID, live_row, correct, congruency, instruction, income_level, gender, handedness) %>%
  drop_na() %>%
  group_by(live_row, congruency, instruction, gender) %>%
  summarize(suc_rate = sum(correct)/n(), .groups = "drop") %>%
  {ggplot(.,aes(x = live_row, y = suc_rate, fill = congruency)) +
    #geom_bar(stat="identity", position=position_dodge()) +
    geom_smooth(method = "loess") +
    ggtitle("Success rate vs. trial no. at instruction and gender") +
    #coord_cartesian(ylim = c(min(.$suc_rate)*.75, 1)) +
    facet_grid(rows = vars(instruction), cols = vars(gender)) +
    theme_minimal()}
```